SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation

### 2018-03-22 paper.tex
preprint

A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup on a Tesla C1060 compared to a serial implementation on a AMD-Opteron processor.
Shi et al.~\cite{Shi:2011aa} paired GPU-based chemical source-term and chemical kinetic Jacobian factorization with two implicit solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU, as compared to a serial Intel i7 930 CPU implementation.
Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential code on an Intel Xeon X5650 CPU.
Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit solver on a Tesla C2050 GPU solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on a Intel Xeon X5650 CPU.
Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics; an order of magnitude was achieved on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
Sewerin and Rigopoulos~\cite{Sewerin20151375} studied an fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU was at best \SI{1.8}{$\times$} slower than the corresponding CPU.
Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.

>>>>

A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup on a Tesla C1060 compared to a serial implementation on a AMD-Opteron processor.
Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU over a serial Intel i7 930 CPU implementation.
Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential code on an Intel Xeon X5650 CPU.
Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit solver on a Tesla C2050 GPU solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on an Intel Xeon X5650 CPU.
Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
